Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-04-28 paper/ms.tex
Cosmetic changes, add surf data

\usepackage[misc]{ifsym}
\usepackage{xcolor}
\usepackage[labelfont=bf]{caption}

@@

\author{Angus Beane$^{1*}$, Lars Hernquist$^1$, Elena D'Onghia$^{2,3}$,
Federico Marinacci$^{4}$, Charlie Conroy$^{1}$, Jia Qi$^{5}$, Laura V.
Sales$^{6}$, Paul Torrey$^{5}$ and Mark Vogelsberger$^{7}$}

@@

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral
  disks\cite{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies \cite{1972MNRAS.157....1L,
  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
  phase of galaxies has conclusively demonstrated that bars should slow their
  rotation over time due to their interaction with dark matter halos
  \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
  2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
  2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
  2009ApJ...697..293D}. We have performed a simulation of a Milky Way-like
  galactic disk hosting a strong bar which includes a state-of-the-art model
  of the interstellar medium and a live dark matter halo. In this simulation
  the bar pattern does not slow down over time, and instead remains at a
  stable, constant rate of rotation. This behavior has been observed in
  previous simulations using more simplified models for the interstellar gas
  but its explanation has remained elusive\cite{1993AA...268...65F,
  2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
  matter halo act in concert to stabilize the bar pattern speed and prevent a
  bar from slowing down or speeding up. This result naturally explains why
  nearly all observed bars rotate rapidly\cite{2011MSAIS..18...23C,
  2015AA...576A.102A, 2019MNRAS.482.1733G} and is especially relevant for our
  understanding of how the Milky Way arrived at its present state.

@@

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig1.pdf}
\caption{Surface density projections in an \Nbody{} only simulation (upper
panels) and a simulation which includes the SMUGGLE model (bottom panels). We
can see that in the \Nbody{} run, the bar grows in length and strength. In the
SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
interstellar medium), the bar remains at approximately the same length and
strength over the course of the simulation. The \Nbody{} model is identical to
the GALAKOS model, discussed in the text.}\label{fig:overview}
\end{figure*}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/fig2.pdf}
\caption{Various bar properties from the \Nbody{} and SMUGGLE runs. The
\textit{upper panel} show the evolution of the pattern speed. As expected, the
bar in the \Nbody{} run slows down due to interactions between the bar and the
dark matter halo. However, the bar in the SMUGGLE run does not slow down and
instead remains at a constant pattern speed. The \textit{middle panel} shows
the evolution of the bar length. In the \Nbody{} case, the bar lengthens. This
occurs because as the pattern speed drops, bar-like orbits at larger radii are
possible. Stars are captured on these orbits, lengthening the bar. This
process does not occur in the SMUGGLE cases since the bar pattern speed is not
decreasing, and therefore the bar length remains constant. The \textit{lower
panel} shows the torque on the bar by different components. The solid lines
show the torque exerted by the halo in both the \Nbody{} and SMUGGLE cases. The
dashed line shows the torque exerted by the gas phase in the SMUGGLE run
(there is no gas in the \Nbody{} run). Details on how these properties
are calculated is given in the Methods section.}\label{fig:prop}
\end{figure}

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig3.pdf}
\caption{The wake excited in the dark matter halo in the \Nbody{} case
(\textit{left panel}) and SMUGGLE case (\textit{right panel}) after
$2.6\,\textrm{Gyr}$. The \textit{left} and \textit{right panels} show a
surface density projection in the $x$-$y$ plane of the dark matter halo after
an axisymmetric average has been subracted. The solid line indicates the
direction of the bar while the dashed line indicates the direction of the halo
wake (both measured by taking the second Fourier component within a sphere of
all material within a radius of $X\,\textrm{kpc}$). The \textit{center panel}
shows the time evolution of the angle difference between the bar and the halo
wake, as measured from the second Fourier component. After the first Gyr, the
angle difference in the SMUGGLE case is smaller than in the \Nbody{} case by
about a factor of two, reflecting how the dark matter halo in the SMUGGLE case
is unable to exert as negative a torque on the bar as in the \Nbody{}
case.}\label{fig:wake}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=9cm]{fig/fig4.pdf}
\caption{The torque on the bar by the gas when the pattern speed of the bar is
kept at a constant value. Only gas within the corotation radius is able to
infall. Since slower bars have larger corotation radii, slower bars experience
a larger net torque than faster bars. The setup of the simulations used here
is identical to the SMUGGLE case discussed earlier and in the Methods section,
except the \Nbody{} disk is rotated as a solid body with a constant angular
velocity.}\label{fig:equil}
\end{figure}

@@

Second, we performed a set of simulations in which we decreased the initial
gas mass of the disk. We essentially find that there is a certain critical gas
fraction (maybe gas surface density?) above which our proposed mechanism
operates. If the gas available is below this threshold, then the pattern speed
will decrease in a similar fashion to standard \Nbody{} simulations.
Importantly, arbitrarily increasing the gas fraction does not cause the
pattern speed to continuously increase. This is consistent with our proposed
equilibrium mechanism and with previous studies of barred galaxies which
varied the gas fraction.\cite{2010ApJ...719.1470V} The exact critical gas
fraction above which our mechanism operates is likely dependent on the
properties of the bar, disk, and dark matter halo -- further work is needed to
characterize how this gas fraction varies with such properties. Details on
this test in which we varied the gas mass of the disk are given in the Methods
section.

The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics.\cite{2021MNRAS.503.2833R,
2021MNRAS.508..926R} Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
2021MNRAS.505.2412C} Third
ed and a fast rotator, of which there is some
observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time (though our work does not
alter expectations for radial mixing induced by spiral
arms\cite{2002MNRAS.336..785S}).

@@

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which are exceptionally
gas-poor to be rotating very slowly. We therefore predict that observations
which target such galaxies (e.g., lenticular barred
galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
exist one such example of a galaxy which is known to be a slow rotator -- UGC
628.\cite{2009AA...499L..25C} This galaxy has been studied in detail by
ref.\cite{2016MNRAS.463.1751C}, who note that it indeed has a low gas
fraction. In general, we predict that all gas-rich spiral galaxies should
rotate quickly while gas-poor spiral galaxies should rotate slowly.

@@

\item [Acknowledgements] We would like to thank Greg L. Bryan, Drummond B.
Fielding, Keith Hawkins, Jason A. S. Hunt, Sarah M. R. Jeffreson, Kathryn V.
Johnston, Julio F. Navarro, David N. Spergel, Martin D. Weinberg, and Yanfei
Zou for helpful discussions. A.B. would like to thank Todd Phillips for
helpful discussions. Some computations in this paper were run on the FASRC
Cannon cluster supported by the FAS Division of Science Research Computing
Group at Harvard University. Resources supporting this work were also provided
by the NASA High-End Computing (HEC) Program through the NASA Advanced
Supercomputing (NAS) Division at Ames Research Center. A.B. was supported by
the Future Investigators in NASA Earth and Space Science and Technology
(FINESST) award number 80NSSC20K1536 during the completion of this work. L.H.
funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding. L.V.S.
funding. P.T. funding. M.V. funding.

@@

\begin{figure}[t]%
\centering
\includegraphics[width=9cm]{fig/fig-vcirc.pdf}
\caption{The circular velocity curve for the \Nbody{} run (blue) and the SMUGGLE
run (orange) compared to observational estimates for the Milky
Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
both runs is marginally larger than the Milky Way's, but still comparable. The
SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
additional mass in the gas phase.}
\label{fig:vcirc}
\end{figure}

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity. We also show the evolution of
the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
in our simulation the atomic and molecular gas surface density and the SFR
surface density is broadly consistent with the expected values for the Milky
Way.\cite{2022ApJ...929L..18E} The discrepancy between $1$ and
$4\,\textrm{kpc}$ is likely due to the fact that the distances to molecular
cloud which underlines this work used a simple kinematic distance based on an
axisymmetric model of the Milky Way,\cite{2017ApJ...834...57M} which is likely
not accurate in the bar region where gas has large non-circular velocities.

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-surf.pdf}
\caption{The time evolution of the molecular gas surface density (\textit{upper}) and
the star formation rate (SFR) surface density (\textit{lower}) at various
times during the simulation. Colored lines indicate the profiles at selected
times during the simulation while the black dashed lines indicate models for
the Milky Way which allow the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$
to vary with metallicity.\cite{2022ApJ...929L..18E} We see that the molecular
gas and SFR surface densities are within an order of magnitude of the Milky
Way's typical values at all times.  We see a sharp decrease in the gas and SFR
surface densities along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
Waiting on atomic gas data from Juergen Kerp.}}
\label{fig:surf}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-A2.pdf}
\caption{The time evolution of the bar strength, measured as the maximum of the second
Fourier component divided by the zeroth Fourier component (a formula is given
in the Methods section). We see that in the \Nbody{} case (blue) the bar
strength increases with time, consistent with previous results showing that
the bar strength increases as bars slow down. In the SMUGGLE case (orange), we
see that the bar strength remains roughly constant, possibly slightly
decreasing with time. This is also consistent with the expected relation
between pattern speed and strength since the bar in this case is not slowing
down.}
\label{fig:strength}
\end{figure}

\begin{figure*}[t!]%
\centering
\includegraphics[width=18cm]{fig/fig-bar_decomp.pdf}
\caption{Demonstration of our bar decomposition procedure based on
ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
projection through the stellar component of the \Nbody{} simulation (disk and
bulge). The \textit{middle panel} shows the component of the disk identified
as being trapped in the bar while the \textit{right panel} shows the component
of the disk identified as not being trapped in the bar. The fact that the
untrapped stars form a roughly axisymmetric structure indicates our bar
decomposition is sufficiently accurate. {\color{red} I will remake this figure
- I ended up losing the files which stored the bar membership, so I just
pulled these screenshots from an earlier presentation.}}
\label{fig:decomp}
\end{figure*}

>>>>

\usepackage[misc]{ifsym}
\usepackage{xcolor}
\usepackage[labelfont=bf]{caption}
\DeclareCaptionLabelSeparator{lsep}{ | } 
\captionsetup{labelsep=lsep}

@@

\author{Angus~Beane$^{1*}$, Lars~Hernquist$^1$, Elena~D'Onghia$^{2,3}$,
Federico~Marinacci$^{4}$, Charlie~Conroy$^{1}$, Jia~Qi$^{5}$,
Laura~V.~Sales$^{6}$, Paul~Torrey$^{5}$ and Mark~Vogelsberger$^{7}$
}

@@

  Elongated bar-like features are ubiquitous in galaxies, occurring at the
  centers of approximately two-thirds of spiral
  disks\cite{2000AJ....119..536E, 2007ApJ...657..790M}.  Due to gravitational
  interactions between the bar and the other components of galaxies, it is
  expected that angular momentum and matter will redistribute over long (Gyr)
  timescales in barred galaxies \cite{1972MNRAS.157....1L,
  1984MNRAS.209..729T, 1985MNRAS.213..451W}. Previous work ignoring the gas
  phase of galaxies has conclusively demonstrated that bars should slow their
  rotation over time due to their interaction with dark matter halos
  \cite{1992ApJ...400...80H, 2000ApJ...543..704D, 2002MNRAS.330...35A,
  2002ApJ...569L..83A, 2003MNRAS.341.1179A, 2003MNRAS.346..251O,
  2005MNRAS.363..991H, 2006ApJ...637..214M, 2007MNRAS.375..460W,
  2009ApJ...697..293D}. We have performed a simulation of a Milky Way-like
  galactic disk hosting a strong bar which includes a state-of-the-art model
  of the interstellar medium and a live dark matter halo. In this simulation
  the bar pattern does not slow down over time, and instead remains at a
  stable, constant rate of rotation. This behavior has been observed in
  previous simulations using more simplified models for the interstellar gas
  but its explanation has remained elusive\cite{1993AA...268...65F,
  2010ApJ...719.1470V}. We propose that the gas phase of the disk and the dark
  matter halo act in concert to stabilize the bar pattern speed and prevent a
  bar from slowing down or speeding up. We find that in a Milky Way-like disk,
  a gas fraction of only about \boldmath$5\%$ is necessary for this mechanism to operate.
  This result naturally explains why nearly all observed bars rotate
  rapidly\cite{2011MSAIS..18...23C, 2015AA...576A.102A, 2019MNRAS.482.1733G}
  and is especially relevant for our understanding of how the Milky Way
  arrived at its present state.

@@

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig1.pdf}
\caption{\textbf{Projected surface densities of simulated barred galaxies with
and without the interstellar medium.} The upper panels show an \Nbody{} only
simulation while the lower panels show a simulation which includes the SMUGGLE
model for the ISM. Columns show different points in time, separated by
$1\,\textrm{Gyr}$. We can see that in the \Nbody{} run, the bar grows in
length and strength. In the SMUGGLE run, the bar remains at approximately the
same length and strength over the course of the simulation. The \Nbody{} model
is identical to the GALAKOS model, discussed in the text.}\label{fig:overview}
\end{figure*}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/fig2.pdf}
\caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
simulations of a barred galaxy.} The \textit{upper panel} shows the evolution
of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
to interactions between the bar and the dark matter halo. However, the bar in
the SMUGGLE run does not slow down and instead remains at a constant pattern
speed. The \textit{middle panel} shows the evolution of the bar length. In the
\Nbody{} case, the bar lengthens. This occurs because as the pattern speed
drops, bar-like orbits at larger radii are possible. Stars are captured on
these orbits, lengthening the bar. This process does not occur in the SMUGGLE
cases since the bar pattern speed is not decreasing, and therefore the bar
length remains constant. The \textit{lower panel} shows the torque on the bar
by different components. The solid lines show the torque exerted by the halo
in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
run). Details on how these properties are calculated is given in the Methods
section.}\label{fig:prop}
\end{figure}

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig3.pdf}
\caption{\textbf{The wake excited in the dark matter halo.} The dark matter
halo wake is shown in the \Nbody{} case (\textit{left panel}) and SMUGGLE case
(\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
\textit{left} and \textit{right panels} show a surface density projection in
the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
been subracted. The solid line indicates the direction of the bar while the
dashed line indicates the direction of the halo wake (both measured by taking
the second Fourier component within a sphere of all material within a radius
of $4\,\textrm{kpc}$). The \textit{center panel} shows the time evolution of
the angle difference between the bar and the halo wake, as measured from the
second Fourier component. After the first Gyr, the angle difference in the
SMUGGLE case is smaller than in the \Nbody{} case by about a factor of two,
reflecting how the dark matter halo in the SMUGGLE case is unable to exert as
negative a torque on the bar as in the \Nbody{} case.}\label{fig:wake}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=9cm]{fig/fig4.pdf}
\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
fixed pattern speed.} Since only gas within the corotation radius is able to
infall and slower bars have larger corotation radii, slower bars experience a
larger net torque than faster bars. The setup of the simulations used here is
identical to the SMUGGLE case discussed earlier and in the Methods section,
except the \Nbody{} disk is rotated as a solid body with a constant angular
velocity.}\label{fig:equil}
\end{figure}

@@

Second, we performed a set of simulations in which we decreased the initial
gas mass of the disk. We essentially find for our disk, our proposed mechanism
operates as long as the gas contributes about $5\%$ of the total mass of the
disk. If the gas available is below this threshold, then the pattern speed
will decrease in a similar fashion to standard \Nbody{} simulations.
Importantly, arbitrarily increasing the gas fraction does not cause the
pattern speed to continuously increase. This is consistent with our proposed
equilibrium mechanism and with previous studies of barred galaxies which
varied the gas fraction.\cite{2010ApJ...719.1470V} The exact critical gas
fraction above which our mechanism operates is likely dependent on the
properties of the bar, disk, and dark matter halo -- further work is needed to
characterize how this gas fraction varies with such properties. Details on
this test in which we varied the gas mass of the disk are given in the Methods
section.

The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics.\cite{2021MNRAS.503.2833R,
2021MNRAS.508..926R} Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar.\cite{2021MNRAS.500.4710C,
2021MNRAS.505.2412C} Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence.\cite{2019MNRAS.490.4740B} And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time (though our work does not
alter expectations for radial mixing induced by spiral
arms\cite{2002MNRAS.336..785S}).

@@

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which are exceptionally
gas-poor to be rotating very slowly. We therefore predict that observations
which target such galaxies (e.g., lenticular barred
galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
exist one such example of a galaxy which is known to be a slow rotator -- the
low surface brightness galaxy UGC 628.\cite{2009AA...499L..25C} This galaxy
has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
indeed has a low gas fraction for galaxies of its type. In general, we predict
that bars in gas-rich spiral galaxies should rotate quickly while bars in
gas-poor spiral galaxies should rotate slowly.

@@

\item [Acknowledgements] We would like to thank Greg~L. Bryan, Drummond~B.
Fielding, Keith Hawkins, Jason~A.~S. Hunt, Sarah~M.~R. Jeffreson, Kathryn~V.
Johnston, Julio~F. Navarro, Joshua~S. Speagle, David~N. Spergel, Martin~D.
Weinberg, and Yanfei Zou for helpful discussions. A.B. would like to thank
Todd Phillips for helpful discussions. Some computations in this paper were
run on the FASRC Cannon cluster supported by the FAS Division of Science
Research Computing Group at Harvard University. Resources supporting this work
were also provided by the NASA High-End Computing (HEC) Program through the
NASA Advanced Supercomputing (NAS) Division at Ames Research Center. A.B. was
supported by the Future Investigators in NASA Earth and Space Science and
Technology (FINESST) award number 80NSSC20K1536 during the completion of this
work. L.H. funding. E.D. funding. F.M. funding. C.C. funding. J.Q. funding.
L.V.S. funding. P.T. funding. M.V. funding.

@@

\begin{figure}[t]%
\centering
\includegraphics[width=9cm]{fig/fig-vcirc.pdf}
\caption{\textbf{The circular velocity curve of our initial setups.} This
curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
compared to observational estimates for the Milky
Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
both runs is marginally larger than the Milky Way's, but still comparable. The
SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
additional mass in the gas phase.}
\label{fig:vcirc}
\end{figure}

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity. We also show the evolution of
the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
in our simulation the atomic and molecular gas surface density and the SFR
surface density is broadly consistent with the expected values for the Milky
Way.\cite{2008AA...487..951K,2022ApJ...929L..18E} The discrepancy between $1$
and $4\,\textrm{kpc}$ in the molecular and SFR surface density is likely due
to the fact that the distances to molecular clouds which underlines this work
used a simple kinematic distance based on an axisymmetric model of the Milky
Way,\cite{2017ApJ...834...57M} which is not accurate in the bar region
where gas has large non-circular velocities.

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-surf.pdf}
\caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
evolution of the atomic gas surface density (\textit{upper}), molecular gas
surface density (\textit{middle}) and the star formation rate (SFR) surface
density (\textit{lower}) at various times during our fiducial simulation.
Colored lines indicate the profiles at selected times during the simulation
while the black dashed lines indicate observations for the atomic
gas\cite{2008AA...487..951K} and black dotted lines indicate a model which
allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
metallicity.\cite{2022ApJ...929L..18E} Molecular gas surface densities were
provided separately (N. Evans, private communication). We see that the
molecular gas and SFR surface densities are within an order of magnitude of
the Milky Way's typical values at all times.  We see a sharp decrease in the
gas and SFR surface densities along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region.}
\label{fig:surf}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-A2.pdf}
\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
the maximum of the second Fourier component divided by the zeroth Fourier
component (a formula is given in the Methods section). We see that in the
\Nbody{} case (blue) the bar strength increases with time, consistent with
previous results showing that the bar strength increases as bars slow down. In
the SMUGGLE case (orange), we see that the bar strength remains roughly
constant, possibly slightly decreasing with time. This is also consistent with
the expected relation between pattern speed and strength since the bar in this
case is not slowing down.}
\label{fig:strength}
\end{figure}

\begin{figure*}[t!]%
\centering
\includegraphics[width=18cm]{fig/fig-bar_decomp.pdf}
\caption{\textbf{Disk decomposition into the bar and unbarred disk.} This
procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
shows a surface density projection through the stellar component of the
\Nbody{} simulation (disk and bulge). The \textit{middle panel} shows the
component of the disk identified as being trapped in the bar while the
\textit{right panel} shows the component of the disk identified as not being
trapped in the bar. The fact that the untrapped stars form a roughly
axisymmetric structure indicates our bar decomposition is sufficiently
accurate. {\color{red} I will remake this figure
- I ended up losing the files which stored the bar membership, so I just
pulled these screenshots from an earlier presentation.}}
\label{fig:decomp}
\end{figure*}
